---
title: "CEDAR KEY, FL INUNDATION"
output:  
  html_document:  
    toc: true  
    toc_float: true  
---

```{r setup, include=FALSE}

library("tidyverse")
library("dplyr")
library("ggplot2")
library("rtide") 
library("scales")
library("BBmisc")
library("lubridate")
```

### INTRODUCTION

This is a document explaining how to use the following code to use the Diver Elevation Equation in relation to the tidal heights collected from Site 3, compensating using NOAA tidal heights. 


### TIDAL HEIGHTS

Use the same NOAA website to download the WATER LEVELS. You can only download the tidal height information for every 30 days. The settings for the tidal heights are:

TIME:GMT  
UNIT: METRIC  
INTERVAL: 1 HOUR  
DATUM: NAVD  

Save the tidal files in another folder, and change the directory. It will be easy to apply the same `lapply` argument to read and combine all of the .csv files. It could be possible that it will not allow you to download the data file per hour because the data is too new, I would wait until that option is available before you download the data. The data files include actual tidal heights and predicted tidal heights. 

```{r tidal_lappy}

setwd("C:/Users/Mel/Desktop/tide_inundation/data/tidal")

#### Verified tidal heights
tidal_temp = list.files(pattern="*.csv")

# Combining all files into a data frame
tidal <- do.call(rbind, lapply(tidal_temp, function(x) read.csv(x, stringsAsFactors = FALSE)))

```

The time is in per hour, so there is no need to round the observations. Time is also in GMT, so it is cohesive with the AIR PRESSURE data. The column names are also updated using `colnames`.

```{r tidal_date}
tidal$date<- paste(tidal$Date, tidal$Time..GMT.)

tidal$date<- as.POSIXct(tidal$date, format="%m/%d/%Y %H:%M")

colnames(tidal)<- c("datex", "timex", "predicted", "prelim", "verified", "date")

```

The following code is to standardize the height data from 0 to 1. The package for the `normalize` argument is `BBmisc` for actual tidal heights. 

```{r tidal_standardize}

tidal$normalization<- normalize(tidal$verified, method= "range", range= c(0,1), margin = 1L, on.constant = "quiet")
```

The following code is to standardize the height data from 0 to 1. The package for the `normalize` argument is `BBmisc` for predicted tidal heights. 
```{r tidal_pre_standardize}
####Predicted tidal 
tidal$pred_normalization<- normalize(tidal$predicted, method= "range", range= c(0,1), margin = 1L, on.constant = "quiet")

```


### SITE 3 BAROMETRIC PRESSURE IN cmH2O

Similar as before, read in the data file, and we are filtering by Site 3, and selecting only the most necessary information for Diver Elevation Equation. 

```{r site_data}
setwd("C:/Users/Mel/Desktop/tide_inundation/data/wq")

site_3<-read.csv("wq.csv", header=T)

#Pressure data are cm
site_3<- site_3 %>% 
  filter(Site==3) %>% 
  select(Date, Pressure, Temperature)

colnames(site_3)<- c("date", "pressure", "temperaure")
```
#### KELVIN
For the Site 3, temperature needs to be converted into Kelvin for the Diver Elevation Equation. The pressure collected from the Diver is cmH2O so we don't need to convert the barometric pressure. 

```{r site_temp_date}
#The Diver Equation requires termperature to be in Kelvin
site_3$kelvin<- (site_3$temperaure + 273.15)
```

### SITE 3 DATE- TIME CORRECTION FOR EST
Site 3, prior to 2019-05-11 06:00, was in time zone EST. After that date/time, it was adjusted to be in UTC timezone. Below is the code to subset and adjust for these times differences. Names of the subsetted data are irrelevant.

```{r wq_date, message=FALSE, warning= FALSE}
site_3$date<- as.POSIXct(site_3$date)

sub.1 <- subset(site_3, date >= as.POSIXct('2017-03-12 02:00:00') &
                  date <= as.POSIXct('2017-11-05 01:00:00'))

sub.1$date<- sub.1$date + 4* 60 *60

sub.2 <- subset(site_3, date >= as.POSIXct('2017-11-05 02:00:00') &
                  date <= as.POSIXct('2018-05-11 05:00:00'))

sub.2$date<- sub.2$date + 5* 60 *60

site<- site_3 %>% 
  filter(!(date <= '2018-05-11 06:00:00'))

site.1<-site_3 %>% 
  filter(date >= '2018-05-11 06:00:00')


site.2<- rbind(sub.1, sub.2)
site_3<- rbind(site.2, site.1)

```


###ELEVATION FORMULA

Elevation Formula as per Diver manual, below are the different variables needed:  

Ph= P0*^- (M*g*H)/(R*T)  

*Ph= atmospheric pressure at elevation height at H*  
*P0= atmospheric pressure at reference height*  
*m= 28.8 *10^-3*  
*g= 9.81 m/s (standard gravity)*  
*R= 8.314 j/mol/k*  
*t= temperature in Kelvin*   
*H= height*  

Water Quality 3 height is -1.422 meters as per project specifications.


```{r}
#Ph is the calculated final value from the inundation equation
site_3$ph<- (site_3$pressure*(-(exp((0.0288*9.81*-1.422)/(8.314*site_3$kelvin)))))

#Standarizing using `normalize` from the package BBmisc
site_3$normalization<- normalize(site_3$ph, method= "range", range= c(0,1), margin = 1L, on.constant = "quiet")
```

###PLOTTING

```{r plot, message= FALSE, warning=FALSE}
ggplot() +
  
  geom_line(data = tidal, aes(x = as.POSIXct(date), y = pred_normalization, color= "Predicted Tidal Height (m)"), size =1.2)  + #<- color in aes() signifies the value of the line, not the physical color, this will generate an appropriate legend
  
  geom_line(data = tidal, aes(x = as.POSIXct(date), y = normalization, color= "Tidal Height (m)"), size =1.2)  +  #<- color in aes() signifies the value of the line, not the physical color, this will generate an appropriate legend
  
  geom_line(data = site_3, aes(x = as.POSIXct(date), y = normalization, color= "Calculated Elevation"), size =1.2)  +  #<- color in aes() signifies the value of the line, not the physical color, this will generate an appropriate legend
   
  scale_x_datetime(name = "January 1-2, 2019 GMT", #<- can be in labs(), also, but fine here, since we need to use scale_x_dateime anyways
                   labels = date_format("%d-%H:%M"),
                   breaks = "4 hours", #<- can adjust the breaks of the x-axis, you can change depending on what your start and end date are
                   limits = c(
                     as.POSIXct("2019-01-01 00:00:00 GMT"), #<- start date and time that can be edited
                     as.POSIXct("2019-01-02 00:00:00 GMT"))) + #<- end date and time that can be edited
  
  #scale_linetype_manual(values = c(1,2,3)) + <- you can use this if you want to change the linetypes
  
  scale_colour_manual(values=c("#0072B2","#D55E00","#CC79A7")) + #<- specifying the colors here, in order as they appear, from a colorblind palette http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
  
  labs(title= "Cedar Key", y= "Tide Height NAVD (m)", color= "Tidal Lines") + #<- labs specify the x-axis, y-axis, and the "color" meaning not the actual colors of the lines but the value we are specifying in the lines, which in this case are Tidal Lines

  theme(legend.position=("bottom"), #<- legend position at the bottom
        panel.border = element_rect(color = "black", size = 1, fill = NA, linetype="solid"), #<- created a border around the figure
        axis.text.x = element_text(angle = 45, hjust = 1)) #<- to make the a-axis ticks 45 degrees
```

